Automated Topic Naming

### 2012-03-16 ese/ese.tex
minor fixes and wording

%Researchers have employed a variety of techniques to extract underlying
%topics that relate to software development artifacts.  Typically, these
%techniques use semi-unsupervised machine-learning algorithms to suggest
%candidate word-lists, topics.  
Software repositories provide a deluge of software artifacts to
analyze. Researchers have attempted to summarize, categorize, and relate these
artifacts by using semi-unsupervised machine-learning
algorithms, such as Latent Dirichlet Allocation (LDA), used for concept and topic analysis to suggest candidate
word-lists or topics that describe and relate software artifacts.
However, these word-lists and topics are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  LDA from commit-log comments recovered
from source control systems. % such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on three large-scale Relational Database Management System (RDBMS) projects: MySQL, PostgreSQL and MaxDB.  
The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels that are relevant to these projects, and provide fresh
insight into their evolving software development activities.
%173 words
\keywords{ Software maintenance \and Repository mining \and Latent Dirichlet allocation \and Topic models}

@@

Concrete applications of \emph{labelled topic extraction} include the annotation of development artifacts with NFR-related tags and the creation of project dashboards.  Annotating software development artifacts with NFR-related tags would allow developers to create detailed directed queries of different artifact kinds that concern the same NFR; for example, a developer could browse the recent history of performance-related bug reports and code check-ins. Project dashboards distill detailed information about a software system into a simpler and more abstract view that summarizes key aspects of the development effort \cite{dashboard}; labelled topic extraction would allow managers to track effort related to specific NFR topics, such as \emph{usability} or \emph{portability}.

@@

% migod
In general, the fruits of mining software artifacts are often project
specific and hard to generalize.  
However, in our previous work we
investigated \emph{topic trends}---that is, topics that recur over
time---where we observed that topic trends often corresponded to
\emph{non-functional requirements} (NFRs)~\cite{Hindle09ICSM}, which is
further emphasized in this paper due to the large numbers of NFR
labelled topics.  
This is encouraging, as NFRs have the property of being cross-domain
and widely applicable. 
In this sense, they are useful abstractions for developer
conversations about different software projects.  
Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying
types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.

@@

This work extends our previous work~\cite{msr2011}. The major extensions in this paper are as follows:
\begin{itemize}
	%\item We added new results on developer activities.
	\item We conducted another case-study on PostgreSQL. 
	\item We used two authors to each annotate the same PostgreSQL topics in order to compare
these annotations. 
	\item We used this new case-study to test \emph{inter-rater
          reliability} (IRR), described in Section \ref{sec:inter-rater}. 
	\item We conducted an analysis of PostgreSQL authors and their association with NFRs and
topics (Section \ref{sec:developers}).
\end{itemize}

@@

Based on this we found that \emph{reliability} and
\emph{usability} worked well for MaxDB in \textsf{exp2} and better in
\textsf{exp3}. 
\textsf{exp1} performed poorly.
MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
but other qualities such as \emph{functionality} did improve. 
For PostgreSQL in \textsf{exp2}, \emph{reliability} and \emph{efficiency} were likewise the most accurate,
while \emph{functionality} remained poor. \emph{Functionality} improved dramatically by \textsf{exp3}.
Our F-measure scores were low and many ROC scores were $0.6$ or less, but our classifier, in most cases,
still performed substantially better than random ($0.5$), even in the
face of heavy class-imbalance for qualities such as \emph{usability} 
and \emph{efficiency}. While there is much room for improvement, we are seeing some
correlation between our quality word lists and relevant topics.

@@

\paragraph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains. 
Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form
of developer comments. 
While we tried to control for application domain variability, OSS
projects investigated were database systems, thus our results might
not generalize well to other domains. Generalizability is enhanced by
the fact that all three projects
were maintained by different developers and different teams using
different development processes.

@@

We validated our topic labelling techniques using multiple experiments. % on unsupervised learners, supervised and multi-label learners. 
We first conducted semi-unsupervised labelling using word-lists. 
Our next approach was supervised, using single-label and multi-label learners. 
Both kinds of learners performed well with average
%area under the 
ROC values between $0.6$ and $0.8$. 
These results were confounded by our low inter-rater reliability score
on the PostgreSQL data-set, suggesting that annotators need careful
and thoughtful training.
With appropriate rigor in annotation, these results, along with the performance of our learners,
demonstrate that labelled topic extraction can be
a promising approach for understanding the occurrence of non-functional requirements in software projects. 
%Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
%We showed that NFRs are often trending topics in commit comments
%, and that NFRs are quite common as developer topics.
%; and that there are efficient methods of automating topic labelling.

>>>>

%Researchers have employed a variety of techniques to extract underlying
%topics that relate to software development artifacts.  Typically, these
%techniques use semi-unsupervised machine-learning algorithms to suggest
%candidate word-lists, topics.  
Software repositories provide a deluge of software artifacts to
analyze. Researchers have attempted to summarize, categorize, and relate these
artifacts by using semi-unsupervised machine-learning
algorithms, such as Latent Dirichlet Allocation (LDA). LDA is used for concept and topic analysis to suggest candidate
word-lists or topics that describe and relate software artifacts.
However, these word-lists and topics are
difficult to  
interpret in the absence of meaningful summary labels.  Current attempts to interpret topics
%topic modeling 
%techniques 
assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  LDA from commit-log comments recovered
from source control systems. % such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on three large-scale Relational Database Management System (RDBMS) projects: MySQL, PostgreSQL and MaxDB.  
The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels that are relevant to these projects, and provide fresh
insight into their evolving software development activities.
%173 words
\keywords{ Software maintenance \and Repository mining \and Latent Dirichlet allocation \and Topic models}

@@

Concrete applications of \emph{labelled topic extraction} include the annotation of development artifacts with non-functional requirements (NFR) keywords, and the creation of project dashboards, which can summarize project activity in a single location.  Annotating software development artifacts with NFR-related tags would allow developers to create detailed directed queries of different artifact kinds that concern the same NFR; for example, a developer could browse the recent history of performance-related bug reports and code check-ins. Project dashboards distill detailed information about a software system into a simpler and more abstract view that summarizes key aspects of the development effort \cite{dashboard}; labelled topic extraction would allow managers to track effort related to specific NFR topics, such as \emph{usability} or \emph{portability}.

@@

% migod
In general, the fruits of mining software artifacts are often project
specific and hard to generalize.  
However, in our previous work we
investigated \emph{topic trends}---that is, topics that recur over
time---where we observed that topic trends often corresponded to
\emph{non-functional requirements} (NFRs)~\cite{Hindle09ICSM}.
%which is further emphasized in this paper due to the large numbers of NFR labelled topics.  
This is encouraging, as NFRs have the property of being cross-domain
and widely applicable. 
In this sense, they are useful abstractions for developer
conversations about different software projects.  
Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying
types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.

@@

This work extends our previous work~\cite{msr2011}. The major extensions in this paper are as follows:
\begin{itemize}
	%\item We added new results on developer activities.
	\item We added a third case study using PostgreSQL, in order to consider another set of data while remaining in the database domain.
	\item We used two authors to each annotate the same PostgreSQL topics in order to compare
these annotations. 
	\item We used this new case-study to test \emph{inter-rater
          reliability} (IRR), described in Section \ref{sec:inter-rater}. 
	\item We conducted an analysis of PostgreSQL authors and their association with NFRs and
topics (Section \ref{sec:developers}).
\end{itemize}

@@

Based on this we found that \emph{reliability} and
\emph{usability} worked well for MaxDB in \textsf{exp2} and better in
\textsf{exp3}. 
\textsf{exp1} performed poorly.
MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
but other qualities such as \emph{functionality} did improve. 
For PostgreSQL in \textsf{exp2}, \emph{reliability} and \emph{efficiency} were likewise the most accurate,
while \emph{functionality} remained poor. \emph{Functionality} improved dramatically by \textsf{exp3}.
Our F-measure scores were low and many ROC scores were $0.6$ or less, but our classifier, in most cases,
still performed substantially better than random ($0.5$), even in the
face of heavy class-imbalance for qualities such as \emph{usability} 
and \emph{efficiency}. While there is much room for improvement, we are seeing some
correlation between our NFR word lists (\textsf{exp2} and \textsf{exp3}) and relevant topics.

@@

\paragraph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
Our data originated from OSS database projects and thus might not be applicable to commercially developed software or software projects in other domains. 
Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form
of developer comments. 
While we tried to control for application domain variability, OSS
projects investigated were database systems, thus our results might
not generalize well to other domains. Generalizability is enhanced by
the fact that all three projects
were maintained by different developers and different teams using
different development processes.

@@

We validated our topic labelling techniques using multiple experiments. % on unsupervised learners, supervised and multi-label learners. 
We first conducted semi-unsupervised labelling using word-lists. 
Our next approach was supervised, using single-label and multi-label learners. 
Both kinds of learners performed noticeably better than random with average
%area under the 
ROC values between $0.6$ and $0.8$. 
These results were confounded by our low inter-rater reliability score
on the PostgreSQL data-set, suggesting that annotators need careful
and thoughtful training.
With appropriate rigor in annotation, these results, along with the performance of our learners,
demonstrate that labelled topic extraction can be
a promising approach for understanding the occurrence of non-functional requirements in software projects. 
%Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
%We showed that NFRs are often trending topics in commit comments
%, and that NFRs are quite common as developer topics.
%; and that there are efficient methods of automating topic labelling.
